---
title: "Peer assignment project 1"
output: html_document
---

```{r}
setwd('/Users/emmasun/Desktop/data science/reproducible research')
library(ggplot2)
library(lattice)
knitr::opts_chunk$set(echo = TRUE)
```
## Loading and preprocessing the data
Load the data
```{r}
rawdata <- read.csv("activity.csv")
```
## What is mean total number of steps taken per day?
###1.Calculate the total number of steps taken per day
```{r}
total_steps <- tapply(rawdata$steps, rawdata$date, FUN=sum, na.rm=TRUE)
```
###2.Make a histogram of the total number of steps taken each day
```{r}
histogram(total_steps, breaks = 50, col = "green", xlab="total number of steps taken per day", ylab = "Frenquency", main = "Histogram of total steps")
```

###3.Calculate and report the mean and median of the total number of steps taken per day

```{r}
mean(total_steps, na.rm=TRUE)
median(total_steps, na.rm=TRUE)
```

## What is the average daily activity pattern?
```{r}
Aver <- aggregate(x=list(steps=rawdata$steps), by=list(interval=rawdata$interval), FUN=mean, na.rm=TRUE)
```
###1.Make a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r}
plot(Aver, type = 'l', col = "green", xlab = "5-minute interval", ylab = "average number of steps taken", main = "Time series")
```

###2.Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
Aver[which.max(Aver$steps),]
```


##Imputing missing values
```{r}
missing <- is.na(rawdata$steps)
```

###1.the total number of missing

```{r}
table(missing)
```

###2.Replacemissing values withmean value of its 5-minute interval

```{r}
subti_value <- function(steps, interval) {
  filled <- NA
  if (!is.na(steps))
    filled <- c(steps)
  else
    filled <- (Aver[Aver$interval==interval, "steps"])
  return(filled)
}
filled_data <- rawdata
filled_data$steps <- mapply(subti_value, filled_data$steps, filled_data$interval)
```

###3.The impact of imputing missing data on the estimates of the total daily number of steps?

```{r}
total_steps <- tapply(filled_data$steps, filled_data$date, FUN=sum)
histogram(total_steps, breaks = 50, col = "green", xlab = "Step/day", ylab = "Frequency", main = "new histogram")
mean(total_steps)
median(total_steps)
```
Answer: Mean and median values are higher after imputing missing data. The reason is that after replacing missing steps values with the mean steps of associated interval value, these 0 vaalues set by default because of those NAs are removed from the histogram of total number of steps taken each day.

##Are there differences in activity patterns between weekdays and weekends?
```{r}
weekday_or_weekend <- function(date) {
  day <- weekdays(date)
  if (day %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday"))
    return("weekday")
  else if (day %in% c("Saturday", "Sunday"))
    return("weekend")
  else
    stop("invalid date")
}
filled_data$date <- as.Date(filled_data$date)
filled_data$day <- sapply(filled_data$date, FUN=weekday_or_weekend)


Aver <- aggregate(steps ~ interval + day, data=filled_data, mean)
ggplot(Aver, aes(interval, steps)) + geom_line() + facet_grid(day ~ .) + xlab("5-minute interval") + ylab("Number of steps")
```
